Factors Affecting Temporal Variations In Vapor Intrusion-Induced Indoor Air Contaminant Concentrations

### 2019-06-04 tex/manuscript.tex
Added implications

There can be a significant role of advective transport in determining observed variability.
For advective transport to play a decisive role, the present results show that there must exist a sub-foundation zone that permits significant flow of soil gas.
In the case of NAS North Island, a relatively permeable sub-foundation soil permitted fairly high variations in building indoor pressure to cause significant fluctuations in contaminant entry rate.
In the case of the ASU house, a permeable gravel sub-base allowed good communication of the indoor air environment with the contents of a preferential contaminant pathway that entered that sub-base, even though the pressure driving force was not nearly as large as that at NAS North Island.
Both situations led to well-documented large variations (two or three orders of magnitude) in indoor air contaminant concentrations.

@@

@@

% PRACTICAL IMPLICATIONS (WEAVE THESE INTO THE TEXT SOMEHOW)
- 24 hr sampling is recommended for a few reasons.
1. Avoids issues with diurnal fluctations in Ae or dP.
2. Variation of a few factors is expected over 24 hrs, while not much else in subsequent days. Thus most short-term variability may be captured in this timeperiod.
- If you are more than (say an order of magnitude) from some regulatory limit in a 24 hr sample, additional sampling at a later date may not be necessary.
- Passive samples can help capture some of the more longterm variations if these are a concern.
- Use manometers to determine dP, if these are relatively large, and if the structure is relatively leaky/the soil is quite permeable, more significant flucations in IACC may be to expected and a "tighter" sampling scheme may be warranted. Of course PP may make even small dP cause significant changes in IACC (which should be screened for).
- PPs should be screened for. Danish study has strategies for this, as well as Pennell. Sampling nearby manholes recommended as these are likely pathways.
- Subslab samples may be missleading. At ASU we saw that alpha > 1 existed, which shows that:
1. Even with "good" subslab sampling, hotspots are easily missed.
2. May mislead that an indoor source is present.
3. Even without a PP there may still be significant spatial variability (Eklund).
4. Holton's tracer study shows that the indoor itself may cause persistent plumes (especially in the deep subslab area).
5. McAlary said something about this. Look it up.
6. May not really be worth the effort.

>>>>

There can be a significant role of advective transport in determining observed variability.
For advective transport to play a decisive role, the present results show that there must exist a sub-foundation zone that permits significant flow of soil gas.
In the case of NAS North Island, a relatively permeable sub-foundation soil permitted fairly high variations in building indoor pressure to cause significant fluctuations in contaminant entry rate.
In the case of the ASU house, a permeable gravel sub-base allowed good communication of the indoor air environment with the contents of a preferential contaminant pathway that entered that sub-base, even though the pressure driving force was not nearly as large as that at NAS North Island.
Both situations led to well-documented large variations (two or three orders of magnitude) in indoor air contaminant concentrations.\par

These findings, as well as those from others reaffirm that screening for preferential pathways should be a routine part of VI investigations.
Sewers have been shown to play significant roles as preferential pathways and sampling nearby manholes for the presence for contaminants is a simple way to determine if a preferential pathway may play a role at the site.
Similar and other suggested screening methods have been described by Nielsen et al.\cite{nielsen_remediation_2017}, Pennell et al.\cite{pennell_sewer_2013}, and McHugh et al.\cite{mchugh_evidence_2017}.\par

Even in the absence of a preferential pathway, monitoring the indoor/outdoor pressure difference at a site may be we wise.
As was observed at North Island NAS, the significant fluctuations in pressurization lead to significant fluctuations in indoor air contaminant concentrations.
Knowing how the pressurization fluctuates may help inform how much temporal variability one may expect at a VI site; larger fluctuations combined with relatively permeable soils and/or leaky buildings lead to more significant temporal variability in indoor air contaminant concentrations, requiring indoor samples to be taken with a finer time-resolution than usual.
Under these circumstances, there may be need for continuous monitoring of indoor air contaminant concentration.\par

@@

This adds to the concerns of overly relying on subslab contaminant samples in VI investigations due to the significant spatial variability of vapor contaminants at VI site like the ASU house and others\cite{chow_concentration_2007,folkes_observed_2009,u.s._environmental_protection_agency_assessment_2015}.
This is further complicated by the fact that the building itself can influence the subslab and create plumes of contaminant vapor\cite{holton_creation_2018}.
Collecting subslab samples can yield important information, but site investigators should be aware of the potential issues and ask themselves if it is worth the effort to collect these samples.\par

@@

This suggests that 24-hour indoor air samples may be the most appropriate choice of sample collection time, avoiding diurnal concerns as well as capturing the variability of a few factors over a day.
Samples collected over longer time-periods, i.e. 48- or 72-hours, are unlikely to yield a different concentration than a 24-hour samples.
For the same reason, multiple samples should be spaced several days or few weeks apart to avoid duplicate samples.
Passive samples could be a cost effective way in conjunction with 24-hour samples to understand the long-term indoor air contaminant concentration trends.
The analysis in Figure \ref{fig:resampling} also implies that if a 24-hour sample gives a indoor air contaminant concentration that deceeds some regulatory limit by an order of magnitude or more, it is unlikely that the indoor air contaminant concentration will exceed this value at a later date; further sampling may not be necessary.\par

\begin{comment}
% PRACTICAL IMPLICATIONS (WEAVE THESE INTO THE TEXT SOMEHOW)
- 24 hr sampling is recommended for a few reasons.
1. Avoids issues with diurnal fluctations in Ae or dP.
2. Variation of a few factors is expected over 24 hrs, while not much else in subsequent days. Thus most short-term variability may be captured in this timeperiod.
- If you are more than (say an order of magnitude) from some regulatory limit in a 24 hr sample, additional sampling at a later date may not be necessary.
- Passive samples can help capture some of the more longterm variations if these are a concern.
- Use manometers to determine dP, if these are relatively large, and if the structure is relatively leaky/the soil is quite permeable, more significant flucations in IACC may be to expected and a "tighter" sampling scheme may be warranted. Of course PP may make even small dP cause significant changes in IACC (which should be screened for).
- PPs should be screened for. Danish study has strategies for this, as well as Pennell. Sampling nearby manholes recommended as these are likely pathways.
- Subslab samples may be missleading. At ASU we saw that alpha > 1 existed, which shows that:
1. Even with "good" subslab sampling, hotspots are easily missed.
2. May mislead that an indoor source is present.
3. Even without a PP there may still be significant spatial variability (Eklund).
4. Holton's tracer study shows that the indoor itself may cause persistent plumes (especially in the deep subslab area).
5. McAlary said something about this. Look it up.
6. May not really be worth the effort.
\end{comment}
